---
title: "Final Project R Reference"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages

### 

##### To clear global environment 
```{r}
rm(list = ls())
```
### require & library
###### Use `require` or `library` to load R packages within your directory.

```{r}
require(here)
require(palmerpenguins)
library(here)
```

### here
###### Use `here` to find project files starting at your working directory.

```{r}
here("data", "ginkgo_data_2021.csv")
```

### Create data frame from .csv
```{r}
gingko = read.csv(here("data", "ginkgo_data_2021.csv"))

summary(gingko)
```


## Data Structures


### c()

***

##### The function `c()` combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type.
* I can’t combine character and numeric types in the same call to c()

Here are two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")
```

Show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
## Typing the name of the vector into the console prints the contents
num_vec
```

```{r}
## The print() function accomplishes the same task:
print(char_vec)
```

### length()

***

##### `Length()` retrieves the length of vectors, lists, and factors. It can be applied to one column or row of a dataset, or any other string of factors, lists, or wherever a vector may appear. 

The following example retrieves the number of values in the vector:

```{r}
length(c(1, 6, 4, 8, 9, 6, 55, 6, 7, 88))
```

The following example returns the number of observations under the "species" column.

```{r}
require(palmerpenguins)
length(penguins$species)
```

### matrix()
       
***

###### To create a matrix: 

Arguments: matrix(data = NA, nrow = 1, ncol = 1, byrow = FALSE,
       dimnames = NULL)
       
* `data` is the data source (ex. `penguins` dataset) 
* `nrow` is the row to select (or number of rows to create)
* `ncol` is column to select (or number of columns to create)
* `byrow = TRUE` orders by row, `FALSE` orders by column
```{r}
mat_1 <- matrix(1:20, 
                ncol = 4,
                )
```



```{r}

```

### nrow()

`nrow` retrieves (or sets) the number of rows from a data frame 

```{r}
nrow(gingko)
```

### ncol()

`ncol()` retrieves (or sets) the number of columns from a data frame

```{r}
ncol(gingko)
```

### dim()

`dim` retrieves the dimensions of a vector or dataframe. `dimnames` sets names for rows and columns during matrix creation. 

```{r}
dim(gingko)
```


## Subsetting

#### Subsetting retrieves specific values, rows, columns, characters, or variables from a dataset (ex. vector, matrix, or data frame). Subsetting can also be used to add or subtract fields to/from data frames, matrices, and vectors

The following operations all perform subsetting. 

### $

```{r}
summary(gingko$max_width)

```

### []

Use `[]` to subset by position

* Can be used on 2 and 3-Dimensional vectors

```{r}
gingko[1,] # select first row of the data frame
gingko[2,3] # select the element in row 2, column 3
head(gingko[, "max_width"]) #select the 3rd column of the data frame


```


### subset

`subset` is a more versatile version of the operation that has the `droplevels` argument in case you need to remove subsets of variables or values from a data set. 

```{r}
penguins = palmerpenguins::penguins
?subset
head(subset(penguins, species == "Adelie"))
```


## Numerical Data Exploration

### summary()

Use `summary` to retrieve these values for a dataset:

* minimum
* 1st Quartile
* Median
* Mean
* 3rd Quartile
* Maximum
```{r}
summary(gingko$max_width)
```

### mean()

Use `mean` to retrieve the mean value for a set of values

```{r}
mean(gingko$max_width)
```

### sd()

Use `sd` to retrieve the standard deviation of a set of values

```{r}
sd(gingko$max_width)
```

## Graphical Data Exploration

### plot()

### Use `plot()` to create scatterplots of variables

```{r}

par(mfrow = c(2, 2))

plot(x = gingko$max_depth, y = gingko$max_width, 
    col = "blue", #  Color 
    pch = 4,  #  Point type 
    cex = .75, #  Point size
    main = "Gingko Leaf Depth and Width", # Header/Title
    xlab = "Maximum Leaf Depth", # X-axis label
    ylab = "Maximum Leaf Width", # Y-axis Label
    xlim = c(1, 150), # X-axis range
    ylim = c(0, 130),  # Y-axis range
)

hist(penguins$flipper_length_mm, main = "Penguin Flipper Lengths", 
     xlab = "Frequency", 
     ylab = "Flipper Length", 
     breaks = 10, # Higher value = greater number of bins
     col = "red", 
     )

boxplot(gingko$petiole_length, main = "Gingko Petiole Lengths",
        ylab = "Petiole Lengths", 
        col = "yellow")

boxplot(formula = max_depth ~ seeds_present,
        data = gingko, 
        main = "Gingko Leaf Depth Conditioned on Petiole Length",
        ylab = "Leaf Depth",
        xlab = "Seed Prescence",
        ls = 2)

```

### hist()

Use `hist()` to create a histogram from given data


```{r}

hist(penguins$flipper_length_mm, main = "Penguin Flipper Lengths", 
     xlab = "Frequency", 
     ylab = "Flipper Length", 
     breaks = 10, # Higher value = greater number of bins
     col = "red", 
     xlim = c(170, 240)
     )

```

### boxplot

Use `boxplot()` to create boxplots of given data. 

```{r}

boxplot(gingko$petiole_length, main = "Gingko Petiole Lengths",
        ylab = "Petiole Lengths", 
        col = "yellow")

boxplot(formula = max_depth ~ seeds_present,
        data = gingko, 
        main = "Gingko Leaf Depth Conditioned on Petiole Length",
        ylab = "Leaf Depth",
        xlab = "Seed Prescence",
        ls = 2)

```

## Distribution Functions

### Normal Distribution

* `dnorm` returns the density function for a normally distributed dataset 
* `pnorm`returns the distribution function for a normally distributed dataset 
* `qnorm` returns the quantile function for a normally distributed dataset 

##### pnorm()

`pnorm()` is a local function in R an is used retrieve the value of the cumulative density function for normally-distributed data using a given `mean`, and standard deviation `sd`

`pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)`

```{r}

````

##### qnorm()

`qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)`

```{r}

````



### Binomial Distribution 

* `dbinom` returns the density function for a normally distributed dataset 
* `pbinom`returns the distribution function for a normally distributed dataset 
* `qbinom` returns the quantile function for a normally distributed dataset 

##### dbinom()

`dbinom(x, size, prob, log = FALSE)`

```{r}

dbinom(0, 6, 2/3)

````


##### pbinom()

`pbinom(q, size, prob, lower.tail = TRUE, log.p = FALSE)`

```{r}
pbinom(4, 6, 2/3)
````

##### qbinom()

`qbinom(p, size, prob, lower.tail = TRUE, log.p = FALSE)`

```{r}

````


